---
title: "tidymodels - bootstrapping for coefficient uncertainty & prediction"
author: "Patrick Ward"
date: "2/25/2023"
output: html_document
---

## Load Packages & Data

```{r}
## packages
library(tidyverse)
library(Lahman)
library(tidymodels)
library(broom)

theme_set(theme_light())

## data
d <- Batting %>%
  filter(yearID >= 2010) %>%
  select(playerID, yearID, AB, R, H) %>%
  group_by(playerID, yearID) %>%
  summarize(across(.cols = everything(),
                   ~sum(.x)),
            .groups = "drop") %>%
  filter(AB >= 200)

d %>%
  head() %>%
  knitr::kable()

```


## EDA

**plot**

```{r}
d %>%
  select(AB:H) %>%
  pivot_longer(cols = everything()) %>%
  ggplot(aes(x = value, fill = name)) +
  geom_histogram(color = "black", 
                 show.legend = FALSE) +
  facet_wrap(~name, scales = "free_x",
             nrow = 2,
             ncol = 2) +
  theme(strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"))
```


**summary statistics**

```{r}
d %>%
  select(AB:H) %>%
  summarize(across(.cols = everything(),
                   list(Avg = mean, Sd = sd))) %>%
  knitr::kable()
```


## Regression Model: R ~ H

```{r}
## Model
fit_lm <- lm(R ~ H, data = d)
tidy(fit_lm)
```


## tidymodels regression with bootstrapping

**create bootstrap folds**

```{r}
### 1000 Bootstrap folds
set.seed(9183)
boot_samples <- bootstraps(d, times = 1000)
boot_samples
```


**Fit a linear model to the bootstrap**

```{r}
fit_boot <-
  boot_samples %>%
  mutate(
    model = map(
      splits,
      ~ lm(R ~ H,
           data = .x)
    ))

fit_boot
```


**Get the coefficients for each of the corresponding bootstraps**

```{r}
boot_coefs <- fit_boot %>%
  mutate(coefs = map(model, tidy))

boot_coefs %>%
  unnest(coefs)
```


## Uncertainty in the Coefficients

**plot uncertainty**

```{r}
boot_coefs %>%
  unnest(coefs) %>%
  select(term, estimate) %>%
  ggplot(aes(x = estimate)) +
  geom_histogram(color = "black",
                 fill = "grey") +
  facet_wrap(~term, scales = "free_x") +
  theme(strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"))
```


**mean and standard deviation of the bootstraps**

```{r}
## bootstrapped coefficient's mean and SD
boot_coefs %>%
  unnest(coefs) %>%
  select(term, estimate) %>%
  group_by(term) %>%
  summarize(across(.cols = estimate,
                   list(mean = mean, sd = sd)))

# check results against linear model
tidy(fit_lm)
```


**exploring the beta coefficients for H**

```{r}
beta_h <- boot_coefs %>%
  unnest(coefs) %>%
  select(term, estimate) %>%
  filter(term == "H")

beta_h %>%
  head()

## 90% Quantile Intervals
quantile(beta_h$estimate, probs = c(0.05, 0.5, 0.95))


## 90% Confidence Intervals
beta_mu <- mean(beta_h$estimate)
beta_se <- sd(beta_h$estimate)

beta_mu
beta_se

beta_mu + qnorm(p = c(0.05, 0.95))*beta_se

## can use the built in function from {tidymodels}
# defaults to a 95% Confidence Interval
int_pctl(boot_coefs, coefs)

# get 90% Confidence Interval
int_pctl(boot_coefs, coefs, alpha = 0.1)
```


## Prediction

**Get the bootstrapped coefficients into a wide data frame**

```{r}
boot_coefs_wide <- boot_coefs %>%
  unnest(coefs) %>%
  select(term, estimate) %>%
  mutate(term = case_when(term == "(Intercept)" ~ "intercept",
                          TRUE ~ term)) %>%
  pivot_wider(names_from = term,
                  values_from = estimate,
              values_fn = 'list') %>%
  unnest(cols = everything())

boot_coefs_wide %>%
  head()

```


**A new batter has 95 hits. How many runs would we predict him to have?**

```{r}
new_H <- 95

new_batter <- boot_coefs_wide %>%
  mutate(pred_R = intercept + H * new_H)

new_batter
```


**Investigate uncertainty for the new batter with 95 hits**

```{r}
## plot the distribution of predictions
new_batter %>%
  ggplot(aes(x = pred_R)) +
  geom_histogram(color = "black",
                 fill = "light grey") +
  geom_vline(aes(xintercept = mean(pred_R)),
             color = "red",
             linetype = "dashed",
             size = 1.4)

## mean and standard deviation of bootstrap predictions
new_batter %>%
  summarize(avg = mean(pred_R),
            SD = sd(pred_R))


## compare to linear model
predict(fit_lm, newdata = data.frame(H = new_H), se = TRUE)
```


**90% Intervals**

* Quantile & Confidence Intervals

```{r}
## get a vector of the predicted runs
pred_runs <- new_batter %>% 
  pull(pred_R)

## 90% Quantile Intervals
quantile(pred_runs, probs = c(0.05, 0.5, 0.95))

## 90% Confidence Interval
mean(pred_runs) + qnorm(p = c(0.025, 0.975)) * sd(pred_runs)

## Compare to 90% confidence intervals from linear model
predict(fit_lm, newdata = data.frame(H = new_H), interval = "confidence", level = 0.90)
```


* Prediction Intervals

```{r}
## For prediction intervals we will need the residual standard error (sigma) for each bootstrap model
boot_sigma <- fit_boot %>%
  mutate(coefs = map(model, glance)) %>%
  unnest(coefs) %>%
  select(id, sigma)

## Get the bootstrap coefficients and the bootstrap id to join the sigma with it
boot_coefs_sigma <- boot_coefs %>%
  unnest(coefs) %>%
  select(id, term, estimate) %>%
  mutate(term = case_when(term == "(Intercept)" ~ "intercept",
                          TRUE ~ term)) %>%
  pivot_wider(names_from = term,
                  values_from = estimate,
              values_fn = 'list') %>%
  unnest(everything()) %>%
  left_join(boot_sigma)

boot_coefs_sigma %>%
  head()

## Now make prediction using a random draw with mean = 0 and sd = sigma for model uncertainty
new_H <- 95

# set seed so that the random draw for model error is replicable
set.seed(476)
new_batter2 <- boot_coefs_sigma %>%
  mutate(pred_R = intercept + H * new_H + rnorm(n = nrow(.), mean = 0, sd = sigma))

new_batter2 %>%
  head()

## plot the distribution of predictions
new_batter2 %>%
  ggplot(aes(x = pred_R)) +
  geom_histogram(color = "black",
                 fill = "light grey") +
  geom_vline(aes(xintercept = mean(pred_R)),
             color = "red",
             linetype = "dashed",
             size = 1.4)

## mean and standard deviation of bootstrap predictions
new_batter2 %>%
  summarize(avg = mean(pred_R),
            SD = sd(pred_R),
            Low_CL90 = avg - 1.68 * SD,
            High_CL90 = avg + 1.68 * SD)

## compare to linear model
predict(fit_lm, newdata = data.frame(H = new_H), interval = "predic", level = 0.9)

```

